---
title: "Worked Example 4 - Parameter Estimation Using Markov Chain Monte Carlo Functions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Worked Example 4 - Parameter Estimation Using Markov Chain Monte Carlo Functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This worked example demonstrates how to use the package to estimate parameters of the yellow fever model from observed epidemiological data within a Bayesian framework using Markov chain Monte Carlo sampling.

We load input data, environmental data and the observed data to be used for comparison in the formats used in previous worked examples (see [Guide 1 - Inputs](https://mrc-ide.github.io/YEP/articles/CGuideAInputs.html) for details of the input format, [Guide 2 - Calculating Parameters From Environmental Data](https://mrc-ide.github.io/YEP/articles/CGuideBParametersEnvironment.html) for details of how environmental data is formatted and used to generate parameter values, and [Worked Example 3 - Generating Epidemiological Data Sets](https://mrc-ide.github.io/YEP/articles/BWorkedExampleCDataGenerate.html) for details of observed data formats):

```{r,eval=FALSE}
library(YEP)
input_data <- readRDS(file = paste(path.package("YEP"), 
                                   "/exdata/input_data_example.Rds", sep = ""))
enviro_data <- read.csv(file = paste(path.package("YEP"), 
                                     "/exdata/enviro_data_example.csv", sep = ""),
                        header = TRUE)
n_env_vars = ncol(enviro_data)-1
obs_sero_data <- read.csv(file = paste(path.package("YEP"), 
                                       "/exdata/obs_sero_data_example.csv", 
                                       sep = ""), header = TRUE)
obs_case_data <- read.csv(file = paste(path.package("YEP"), 
                                       "/exdata/obs_case_data_example.csv", 
                                       sep = ""), header = TRUE)
obs_outbreak_data <- NULL
# Values of environmental coefficients to use as initial parameter values
enviro_coeffs_ini <- as.numeric(read.csv(file=paste(path.package("YEP"),
                        "/exdata/enviro_coeffs_example.csv", sep = ""), header = TRUE)) 
```

We set additional inputs as follows. The parameters varied in the chain are the natural logarithm values of the actual parameters of interest; the inputs log_params_min, log_params_max and log_params_ini are set accordingly. The maximum, minimum and initial values of the varied parameters must be entered into log_params_min, log_params_max and log_params_ini in the following order:

-Values of spillover FOI or the coefficients used to calculate spillover FOI from environmental data  
-Values of R0 or the coefficients used to calculate R0 from environmental data (if R0 is to be varied)  
-Vaccine efficacy (if varied)  
-Severe non-fatal case reporting probability (if varied)  
-Fatal case reporting probability (if varied)

The maximum and minimum values are not all necessarily used; if prior_type is set to "flat", then they are hard limits for all parameters. If prior_type is set to "zero" or "norm", only the limits for p_rep_severe and p_rep_death are used.

```{r,eval=FALSE}
# Initial log parameter values
log_params_ini <- c(log(enviro_coeffs_ini), log(c(0.1,0.2))) 
# Prefix to be used for .csv output files; a new file is created every 10,000 
# iterations with incremental number suffixes (Chain0.csv, Chain1.csv, etc.)
filename_prefix <- "Chain" 
#Number of iterations to run
Niter <- 1 
# Flag indicating how to set initial population immunity level
mode_start <- 1 
# Type of prior likelihood calculation to use
prior_settings <- list(type="zero")
# Time increment in days
time_inc <- 5.0 
# Number of repetitions for which to run model and obtain average outputs
n_reps <- 5 
# Probability of an infection causing severe symptoms
p_severe_inf = 0.12
# Probability of an infection with severe symptoms causing death
p_death_severe_inf = 0.39
# Additional values: severe and fatal case reporting probability are set to NULL due to being estimated
# as variables; vaccine efficacy and Brazil spillover FOI multiplier set to 1.0 as constants.
add_values=list(p_rep_severe = NA, p_rep_death = NA, vaccine_efficacy = 1.0, m_FOI_Brazil = 1.0)
# True/false flag indicating whether or not to run model in deterministic mode
# (so that binomial calculations give average instead of randomized output)
deterministic = FALSE
# Variable to set different modes for running on multiple processors simultaneously;
# here set to "none" so that parallel processing is not used
mode_parallel=FALSE
```

We then run the MCMC() function. Output data recorded at each iteration is saved to an output file every 10 iterations. Likelihood is calculated as described in Guide 4 [Link TBA].

```{r,eval=FALSE}
MCMC(log_params_ini, input_data, obs_sero_data, obs_case_data, filename_prefix, 
     Niter, mode_start, prior_settings, time_inc, n_reps, enviro_data, p_severe_inf, 
     p_death_severe_inf, add_values, deterministic,mode_parallel,NULL)
```

The saved output of MCMC() takes the form of a data frame of values of the current likelihood (posterior_current), likelihood calculated from proposed parameter values (posterior_prop), current and proposed parameter values, a flag indicating whether the proposed parameter values were accepted (set to 1 if they were accepted, 0 if they were not), and the chain covariance.

```{r,eval=FALSE}
chain_example <- read.csv(file = paste(path.package("YEP"), 
                                       "/exdata/Chain_example.csv", sep = ""),
                          header = TRUE)
head(chain_example, 1)
```
