---
title: "Worked Example 1 - Single Model Run"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Worked Example 1 - Single Model Run}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This worked example demonstrates how to set up and carry out a single run of the model.

The model of yellow fever used in YEP is an age-stratified SEIRV model in which the number of people in each age group in groups S (susceptible), E (exposed), I (infectious), R (recovered) and V (vaccinated) is computed at each time point, with changes in population also incorporated. The number of new infections (individuals moving from group S to group E) at each time point is calculated stochastically using the total force of infection (FOI) as a binomial probability. A more detailed mathematical description of the model is given in the supplementary information to [Assessing yellow fever outbreak potential and implications for vaccine strategy](https://doi.org/10.1371/journal.pgph.0003781) in PLOS GPH by Fraser et al.

Population and vaccination data is stored in a list object as described in [Guide 1 - Inputs](https://mrc-ide.github.io/YEP/articles/CGuideAInputs.html). Here we load an example data set for multiple regions and extract the vaccination and population data for the first two included regions:

```{r,eval=FALSE}
library(YEP)
input_data <- readRDS(file = paste(path.package("YEP"), 
                                   "/exdata/input_data_example.Rds", sep = ""))
vacc_data <- input_data$vacc_data[c(1:2), , ]
pop_data <- input_data$pop_data[c(1:2), , ]  
```

We set the other input parameters to the functions which run the model as follows. Here we set the spillover FOI and basic reproduction number directly; they may alternatively be calculated from environmental covariates as described in [Guide 2 - Calculating Parameters From Environmental Data](https://mrc-ide.github.io/YEP/articles/CGuideBParametersEnvironment.html). The mode_start variable is used to set the initial conditions in one of 3 ways. The n_particles variable is used to run the model multiple times for the same inputs; the results can be compared and/or combined to test and/or average out the effects of the stochasticity of the model. The n_threads variable is used to set the number of particles run in parallel (subject to the limitations of the computer being used). 

```{r,eval=FALSE}
# Force of infection due to spillover from sylvatic reservoir (here fixed over time with one value per region; must be a 2-D array with dimensions (number of regions, number of time points))
FOI_spillover <- array(c(1.0e-8, 1.0e-8),dim=c(2,1))
# Basic reproduction number for urban spread of infection (here fixed over time with one value per region; must be a 2-D array with dimensions (number of regions, number of time points))
R0 <- array(c(1.2, 1.2),dim=c(2,1))
# Year(s) for which data is to be output
years_data <- c(1995:2000)         
# First year in population/vaccination data (here taken from the input data set; 
# if the population and vaccination data is set manually, this similarly needs 
# to be set manually)
year0 <- input_data$years_labels[1]
# Proportional vaccine efficacy (from 0 to 1); probability of vaccination 
# successfully providing immunity (i.e. moving vaccinated individual to group V)
vaccine_efficacy <- 1.0      
# Time increment in days to use in model (should be 1.0, 2.5 or 5.0 days)
time_inc <- 1.0     
# Type of data to output
#   If mode_out = 1, SEIRVC + FOI for all regions, steps and ages at every time point in years_data
#   If mode_out = 2 annual totals of SEIR, R and V by region and age for calculating seroprevalence
#   If mode_out = 3, annual total new infections (C_annual) by region combined across age groups
#   If mode_out = 4, annual total new infections (C_annual) not combined by age
#   If mode_out = 5, total new infections (C) combined by age at every time point in years_data
mode_out <- 1
# Flag indicating how to set initial population immunity level in addition to
# vaccination
# If mode_start = 0, only vaccinated individuals are immune (R = 0)
# If mode_start = 1, shift some non-vaccinated individuals into recovered to 
# give herd immunity (stratified by age)
# If mode_start = 2, use SEIRV input in list from previous run(s), set using
# start_SEIRV variable
mode_start <- 1           
# SEIRV data from end of a previous run to use as input (if mode_start = 2; not 
# used here and hence set to NULL)
start_SEIRV <- NULL   
# Type of time variation of FOI_spillover and R0; here set to 0 for constant values
mode_time <- 0
# Number of particles to run
n_particles <- 10     
# Number of threads to run
n_threads <- 1   
# True/false flag indicating whether or not to run model in deterministic mode
# (so that binomial calculations give average instead of randomized output)
deterministic <- FALSE
# Optional random seed value to set for stochastic mode; set to NULL if unused
seed <- 1
```

We use the Model_Run() function to run the model for the inputs above:

```{r,eval=FALSE}
test1 <- Model_Run(FOI_spillover, R0, vacc_data, pop_data, years_data, year0,
                   vaccine_efficacy, time_inc, mode_out, mode_start,
                   start_SEIRV, mode_time, n_particles, n_threads, deterministic, seed)
```

The output of Model_Run(), when output_type is set to "full", includes, for each particle, date information, the total force of infection FOI_total at each data point, the numbers of people in groups S, E, I, R and V at each time point organized by age group, and the number of newly infectious people (C) at each time point organized by age group. We use the plot_model function from the YEPaux package to display values of S, R and V (added together across all age groups) over time, with error bars to show the variation between particles caused by stochasticity if relevant.

```{r,eval=FALSE}
#TBC
library(YEPaux)
plot1 <- plot_model_output(test1)
print(plot1)
```

The model can be run in deterministic mode by setting the 'deterministic' flag to TRUE. In this mode, the number of new infections at each time point is calculated by multiplying the number of susceptible individuals by the total force of infection, rather than via sampling a binomial distribution.

```{r,eval=FALSE}
test2 <- Model_Run(FOI_spillover, R0, vacc_data, pop_data, years_data, year0,
                   vaccine_efficacy, time_inc, mode_out, mode_start,
                   start_SEIRV, mode_time, n_particles, n_threads, 
                   deterministic=TRUE, seed=NULL)
#TBC
plot2 <- plot_model_output(test2)
print(plot2)
```
